---
title: "lynall07"
output: html_document
---

## Contents:
- Stepwise models to regress out sex, BMI, age, tobacco, recent infection, using whole dataset
- Forced two-way clustering of MDD residual cell counts
- Plots in Figures 4E, 4F
- Plots in Figure S6

```{r}

library(ggplot2)
library(dplyr)
library(ggsignif)
library(tidyr)
library(magrittr)
library(reshape2)
library(scales)
library(here)

# Load full dataset
load(file=paste(here::here("res/"),"Apollo_qm_combi.RData",sep=""))

# Cell counts and labels
choice_qset <- c("q_rbc","q_platelet","q_baso","q_eosin","q_neut","q_mono_class","q_mono_int","q_mono_nonclass","q_cd4","q_cd8","q_b","q_nkt","q_nk16","q_nk56")
labels_qset <- c(q_rbc="RBCs",q_platelet="Platelets",q_baso="Basophils",q_eosin="Eosinophils",q_neut="Neutrophils",q_mono_class="Mono (classical)",q_mono_int="Mono (int)",q_mono_nonclass="Mono (non-class)",q_cd4="CD4+ T",q_cd8="CD8+ T",q_b="B cells",q_nkt="NKT cells",q_nk16="CD16(hi) NK",q_nk56="CD56(hi) NK")

```

## Generating residual cell counts.

Steps for each cell count are:
(a) box-cox transform
(b) stepwise regression (initial model including AGE, BMI, SEX, CENTRE, TOBACCO, INFECTIONS)
(c) clustering of residuals from model selected in (b)
New 'residual' cell counts are listed in choice_reg, each of the 14 named "reg_xxx"

Debatching and stepwise modeling require initial imputing of missing BMI, tobacco use and infection status. Imputed values are not used for statistical comparisons of clinical groups or clusters.
```{r}

choice_debatch <- choice_qset

sapply(qm_combi[,c("bmi_calculated","tobacco_yn","infections")],function(x) sum(is.na(x))) 

# 8 bmis are missing, 4 tobacco use missing, 6 infection status missing

# BMI - impute median
qm_combi$bmi_imputed <- ifelse(is.na(qm_combi$bmi_calculated), median(qm_combi$bmi_calculated, na.rm=T), qm_combi$bmi_calculated)

# Tobacco - impute mode (no tobacco)
qm_combi$tobacco_imputed <- qm_combi$tobacco_yn
qm_combi$tobacco_imputed %<>% forcats::fct_explicit_na()
qm_combi$tobacco_imputed %<>% forcats::fct_collapse(no=c("no","(Missing)"))
qm_combi$tobacco_imputed

# Infection - impute mode (no infection)
qm_combi$infections_imputed <- qm_combi$infections
qm_combi$infections_imputed[is.na(qm_combi$infections_imputed)] <- "no"

# Box-cox transform
prepro <- caret::preProcess(qm_combi[,choice_debatch],method="BoxCox",na.remove=TRUE) 
pppls <- predict(prepro,qm_combi[,choice_debatch])

# Debatch / regress out variables
set.seed(111)
steps <- list()
steps <- sapply(colnames(pppls), function(x){
  print(x)
  llm <- lm(pppls[,x] ~ qm_combi$sex + qm_combi$bmi_imputed + qm_combi$age_at_bloods + qm_combi$infections_imputed + qm_combi$tobacco_imputed)
  step <- MASS::stepAIC(llm, direction="both")
  print(summary(step))
  return(step)
})

residuals <- as.data.frame(sapply(names(steps), function(x){steps[[x]]$residuals} ))

# Rename and merge to m
colnames(residuals) <- paste("reg",colnames(residuals),sep="_") 
residuals$name <- qm_combi$name
qm_combi <- left_join(qm_combi,residuals,by="name")
choice_reg <- paste("reg",choice_qset,sep="_")

```

## PCA of residuals
```{r}

mdd <- qm_combi[qm_combi$disorder=="MDD",]
tmp <- mdd[,choice_reg]
colnames(tmp) <- labels_qset
pca.out <- prcomp(tmp,center=T,scale=T)

# Flip axes for consistency with previous PCA plots
pca.out$rotation[,c(1,2)] %<>% "*"(-1)
pca.out$x[,c(1,2)] %<>% "*"(-1)

library(reshape2)
eigen <- data.frame(pca.out$rotation[,1:2])
colnames(eigen) = c("PC1","PC2")
eigen$count <- rownames(eigen)

# New sorting order
desired_order <- rev(c("Neutrophils","Eosinophils","Basophils","Mono (non-class)","Mono (int)","Mono (classical)", "CD4+ T" ,"CD8+ T" ,"CD16(hi) NK","CD56(hi) NK","NKT cells","B cells" ,"RBCs","Platelets"))            

# Re-order the levels
eigen$count <- factor( as.character(eigen$count), levels=desired_order )
# Re-order the data.frame
eigen <- eigen[order(eigen$count),]

ggplot(melt(eigen, value.name="Eigenvector"),aes(x=count,y=Eigenvector, fill=variable)) + geom_bar(stat="identity") + coord_flip() + scale_x_discrete(labels=labels_qset) + facet_grid(~variable, scales="free") + xlab("Cell count") + scale_fill_discrete(name = "Component") + guides(fill=FALSE) + theme_bw()
ggsave(file=paste(here::here("pics/"),"apollo_mdd_pca_pca1and2_RESIDUALS.pdf",sep=""),width=4.5,height=3.5)

```

PLS-R: cell count residuals predict depression symptom scores
```{r}

# Function for multivariate permutation testing modified from getMethod("opls”,"matrix”)
plsr_multivariate_perm <- function(opl, permI){
  modSumVc <- colnames(opl@summaryDF)
  permMN <- matrix(0, nrow = 1 + permI, ncol = length(modSumVc), 
                   dimnames = list(NULL, modSumVc))
  perSimVn <- numeric(1 + permI)
  perSimVn[1] <- 1
  permMN[1, ] <- as.matrix(opl@summaryDF)
  for (k in 1:permI) {
    yVcn <- drop(opl@suppLs[["yMCN"]]) # Matrix of outcomes
    yPerVcn <- yVcn[sample(nrow(yVcn)),] # Permute the rows. Will maintain the interdependencies.
    yPerMCN <- as.matrix(yPerVcn)
    rownames(yPerMCN) <- NULL 
    predI <- opl@summaryDF$pre 
    perOpl <- opls(yPerMCN,as.matrix(full[,yin]), plotL=FALSE, predI=predI, printL=FALSE) 
    permMN[1 + k, ] <- as.matrix(perOpl@summaryDF)
  }
  perPvaVn <- c(pR2Y = (1 + length(which(permMN[-1, "R2Y(cum)"] >= permMN[1, "R2Y(cum)"])))/(nrow(permMN) - 1), # number of perm R2Y(cum) above actual R2Y(cum)
                pQ2 = (1 + length(which(permMN[-1, "Q2(cum)"] >= permMN[1, "Q2(cum)"])))/(nrow(permMN) - 1)) # number of Q2(cum) above actual R2Y(cum)
  opl@summaryDF[, "pR2Y"] <- perPvaVn["pR2Y"]
  opl@summaryDF[, "pQ2"] <- perPvaVn["pQ2"]
  opl@suppLs[["permMN"]] <- permMN
  return(opl)
}


library(ropls)
yin <- c("ham_17_score","bdi_total_score","shaps_score_calc","chalder_fatigue_score")
labels_yin <- c("HAM Depression","Beck Depression","SHAPS anhedonia","Chalder fatigue")
full <- qm_combi[complete.cases(qm_combi[,c(yin,choice_reg)]) & qm_combi$disorder=="MDD", ]
rownames(full) <- full$name

set.seed(111)
m.pls <- opls(full[,choice_reg],as.matrix(full[,yin]), plotL=FALSE) # Single component model
m.pls.with.perm <- plsr_multivariate_perm(m.pls, permI=5000)
m.pls.with.perm
# PLS
# Apollo

m.pls@modelDF

library(boot)
boot.out <- list()
  getweight <- function(data,ind){
    x <- data[ind,] # Allow bootstrap to select subset
    m.pls <- opls(x[,choice_reg],as.matrix(x[,yin]), permI=0, predI=1, printL=FALSE, plotL=FALSE)
    tmp <- getWeightMN(m.pls)
    return(tmp)
  }
  
  set.seed(222)
  boot.out <- boot(full,getweight,R=1000) 

boot.z <- boot.out$t0 / apply(boot.out$t,2,sd) 
boot.z[rev(order(boot.z)),]

# Force two-component model for plotting purposes
m.pls.two <- opls(full[,choice_reg],as.matrix(full[,yin]), permI=1,predI=2, plotL=FALSE)

# Scores plot 
ggplot(data=as.data.frame(m.pls.two@scoreMN),aes(x=p1,y=p2)) + geom_point(aes(colour=full$disorder)) + ggtitle("")  + theme_bw() + geom_hline(yintercept =0) + geom_vline(xintercept=0) + stat_ellipse() + xlab("Component 1") + ylab("Component 2") # + expand_limits(x=-7.5)
# +geom_text(aes(label=res,size=1,hjust=0,vjust=0))
ggsave(file=paste(here::here("pics/"),"apollo_pls_matrix_symptoms_scores_mdd_REG.pdf",sep=""), width=7, height=5) 

# Weights plot
tmp <- as.data.frame(m.pls.two@weightStarMN)
tmp$component <- "predictor"
tmp$boot.z <- boot.z
tmp2 <- as.data.frame(m.pls.two@cMN)
tmp2$component <- "response"
tmp2$boot.z <- 999 # Set >3 to force printing
(weights_combi <- bind_rows(tmp, tmp2))
weights_combi$label <- c(labels_qset,labels_yin)

library(ggrepel)
fig2b <- ggplot(data=weights_combi,aes(x=p1,y=p2, colour=component, label=label)) + geom_point() + ggtitle("")  + theme_bw() + geom_hline(yintercept =0) + geom_vline(xintercept=0) + xlab("Component 1") + ylab("Component 2") + guides(color=guide_legend(title="Variables")) + geom_label_repel(data=dplyr::filter(weights_combi,boot.z>3),aes(x=p1,y=p2, colour=component)) + expand_limits(y=c(0.45), x=c(0.55)) + scale_color_manual(values=c("dark green","purple")) + theme(legend.position = c(0.85,0.15),legend.background = element_rect(color = "black"), legend.title = element_blank()) 
ggsave(fig2b, file=paste(here::here("pics/"),"apollo_pls_matrix_symptoms_weights_mdd_REG.pdf",sep=""),width=6, height=5.4) 

```

## Consensus clustering of residuals
```{r}
# First, simply cluster all MDD participants to check model selected
library(mclust)
set.seed(222)
mod1 <- Mclust(mdd[,c(choice_reg)],G=2, verbose=TRUE) 
summary(mod1, parameters = FALSE)  # n=106 and n=100. EEI model.

# Now, do with 90% subsampling
library(clue)
nr <- NROW(mdd)
set.seed(222)
tmp2 <- replicate(5000, expr = {
  out <- Mclust(mdd[sample(nr, 0.9 * nr, replace = FALSE),choice_reg],modelNames = "EEI", G=2, verbose = FALSE) # Forcing G to be 2
  as.cl_partition(cl_predict(out, mdd[,choice_reg], "memberships"))
},simplify = FALSE) # subsample
tmp3 <- cl_ensemble(list=tmp2)
set.seed(222)
tmp4 <- cl_consensus(tmp3) 
meta_two_way_resid <- cl_class_ids(tmp4); table(meta_two_way_resid) # 102 and 104

mdd$meta_two_way_resid <- factor(meta_two_way_resid)

# Load the original 4-way and forced 2-way clustering of absolute cell counts
load(paste(here::here("res/"),"Apollo_mdd_force_two_way.Rdata",sep=""))
mdd$meta_two_way <- factor(meta_two_way)
load(paste(here::here("res/"),"Apollo_mdd_meta_subsample.R",sep=""))
mdd$meta_subsample <- factor(meta_subsample)
mdd %>% group_by(meta_two_way_resid) %>% summarise_at(vars(ham_17_score),median) # 16 and 18

mdd$meta_two_way_resid %<>% forcats::fct_recode("C0"="2","C1"="1") 
mdd$meta_two_way_resid %<>% factor(levels=c("C0","C1"), ordered=TRUE)
meta_two_way_resid <- mdd$meta_two_way_resid

```

Function to plot cluster features
```{r}
make_mini_cluster_plot_force2 <- function(data=m, basedir=here::here("res/"), picpath=here::here("pics/"), cluster_fx=choice_qset, choice_qset=choice_qset, reg=NULL, cols=clustercols, levels=levels, suffix=""){
  
  vjust <- 0.5
  step_increase <- 0.12
  
  if (reg=="reg"){ # If using residuals counts rather than absolute counts
    data[,choice_qset] <- NULL 
    colnames(data) <- sapply(colnames(data), function(x){gsub("reg_","",x)}) 
    cluster_fx <- choice_qset
    rm(choice_qset) 
  }
  
  print(paste("Number of samples finally included in clustering = ",dim(data)[1],sep=""))
  
  set.seed(111)
  
  nclust <- length(levels(data$cluster_id))
  
  ching <- function(data=NULL, dependent=NULL){
    chi1 <- chisq.test(t(table(data[,dependent], data$cluster_id)),simulate.p.value = TRUE) # Cluster ID as rows (=populations to be compared)
    print(paste("Chis for variable:",dependent))
    print(t(table(droplevels(data[,dependent]), data$cluster_id)))
    return(chi1)
  }
  
  categoricals <- c("centre","infections","lynall_inflamm_yn","tobacco_yn","alcohol_yn","cannabis_yn","current_antidepress_yn", "sex")
  
  print("testing chi for...")
  print(categoricals)
  
  all_ching <- lapply(categoricals, function(x) {ching(data=data,dependent=x)})
  names(all_ching) <- categoricals 
  print("2 clusters, chi2 results are:")
  print(all_ching)
  
  # Kruskall-Wallis
  krusking <- function(data=NULL, dependent=NULL){
    kruk <- PMCMRplus::kruskalTest(get(dependent) ~ cluster_id, data=data)
    conov <- PMCMRplus::kwAllPairsConoverTest(get(dependent) ~ cluster_id, data=data, p.adjust.method="BH")
    conov <- conov$p.value 
    if(kruk$p.value<=0.05){ 
      conov <- conov
    } else {
      conov[,] <- NA
    }
    kruker <- list("p.value"=kruk$p.value, "conov"=conov)
    print(paste("Kruskal for variable:",dependent))
    print(kruker$p.value)
    return(kruker)
  }
  numerics <- c(cluster_fx,"bmi_calculated","age_at_bloods","crp_macs","p_il_6_mean","ham_17_score","ctq_total_score","chalder_fatigue_score","bdi_total_score","shaps_score_calc","t_anxiety_score","s_anxiety_score","antidepress_failed_75_less", "leq_z")
  
  print(numerics)
  all_krusking <- lapply(numerics, function(x) {krusking(data=data,dependent=x)})
  print(length(all_krusking))
  names(all_krusking) <- numerics 
  print(names(all_krusking))
  
  # Plots
  
  p_sex <- ggplot(data=data, aes(sex, fill = cluster_id, label=..count..)) + geom_bar(position = "fill") + 
    scale_y_continuous(labels = scales::percent) + xlab("") + ylab("") + 
    ggtitle("Sex\n") + scale_fill_manual(values=cols) + guides(fill=FALSE) + 
    theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + 
    geom_text(stat="count",position="fill",vjust=1)
  
  p_infec <- ggplot(data=data[!is.na(data$infections),], aes(droplevels(infections), fill = cluster_id, label=..count..)) + geom_bar(position = "fill") + 
    scale_y_continuous(labels = scales::percent) + xlab("") + ylab("") + 
    ggtitle("Infection\nstatus\n") + scale_fill_manual(values=cols) + guides(fill=FALSE) + 
    theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + 
    geom_text(stat="count",position="fill",vjust=1)
  
  p_inflamm <- ggplot(data=data[!is.na(data$lynall_inflamm_yn),], aes(droplevels(lynall_inflamm_yn), fill = cluster_id, label=..count..)) + geom_bar(position = "fill") + 
    scale_y_continuous(labels = scales::percent) + xlab("") + ylab("") + 
    ggtitle("Minor\ninflammatory\ndisease") + scale_fill_manual(values=cols) + guides(fill=FALSE) + 
    theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + 
    geom_text(stat="count",position="fill",vjust=1)
  
  p_tobacco <- ggplot(data=data[!is.na(data$tobacco_yn),], aes(tobacco_yn, fill = cluster_id, label=..count..)) + geom_bar(position = "fill") + 
    scale_y_continuous(labels = scales::percent) + xlab("") + ylab("") + 
    ggtitle("Current\ntobacco use\n") + scale_fill_manual(values=cols) + guides(fill=FALSE) + 
    theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + 
    geom_text(stat="count",position="fill",vjust=1)
  
  p_alcohol <- ggplot(data=data[!is.na(data$alcohol_yn),], aes(alcohol_yn, fill = cluster_id, label=..count..)) + geom_bar(position = "fill") + 
    scale_y_continuous(labels = scales::percent) + xlab("") + ylab("") + 
    ggtitle("Current\nalcohol use\n") + scale_fill_manual(values=cols) + guides(fill=FALSE) + 
    theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + 
    geom_text(stat="count",position="fill",vjust=1)
  
  p_cannabis <- ggplot(data=data[!is.na(data$cannabis_yn),], aes(cannabis_yn, fill = cluster_id, label=..count..)) + geom_bar(position = "fill") + 
    scale_y_continuous(labels = scales::percent) + xlab("") + ylab("") + 
    ggtitle("Current\ncannabis use\n") + scale_fill_manual(values=cols) + guides(fill=FALSE) + 
    theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + 
    geom_text(stat="count",position="fill",vjust=1)
  
  # Percentage currently using antidepressant
  p_propAD <- ggplot(data=data[!is.na(data$current_antidepress_yn),], aes(current_antidepress_yn, fill = cluster_id, label=..count..)) + geom_bar(position = "fill") + 
    scale_y_continuous(labels = scales::percent) + xlab("") + ylab("") + 
    ggtitle("Current anti-\ndepressant\ntreatment\n") + scale_fill_manual(values=cols) + 
    theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + 
    guides(fill=FALSE) + 
    geom_text(stat="count",position="fill",vjust=1)
  
  p_age_onset <- ggplot(data=data, aes(x=cluster_id, y=age_onset_lynall)) + xlab("") + ylab("Years") + 
    ggtitle("Age of\nonset") + geom_violin(adjust=2) + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + 
    expand_limits(y=c(0)) +
    guides(fill=FALSE) + theme_bw()
  
  p_time_since_onset <- ggplot(data=data, aes(x=cluster_id, y=time_since_onset)) + xlab("") + ylab("Years") + 
    ggtitle("Time since\nonset") + geom_violin(adjust=2) + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + 
    expand_limits(y=c(0)) +
    guides(fill=FALSE) + theme_bw()
  
  p_mc <- ggplot(data=data, aes(x=cluster_id, y=q_mono_class)) + xlab("") + ylab("") + 
    ggtitle("Classical\nmonocytes") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols)+ guides(fill=FALSE) + expand_limits(y=0) + theme_bw() 
  
  p_mi <- ggplot(data=data, aes(x=cluster_id, y=q_mono_int)) + xlab("") + ylab("") + 
    ggtitle("Intermediate\nmonocytes") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0)+ theme_bw() 
  
  p_mnc <- ggplot(data=data, aes(x=cluster_id, y=q_mono_nonclass)) + xlab("") + ylab("") + 
    ggtitle("Non-classical\nmonocytes") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0)+ theme_bw() 
  
  p_b <- ggplot(data=data, aes(x=cluster_id, y=q_b)) + xlab("") + ylab("") + 
    ggtitle("B cells") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + expand_limits(y=0) + theme_bw() 
  
  p_cd4 <- ggplot(data=data, aes(x=cluster_id, y=q_cd4)) + xlab("") + ylab("") + 
    ggtitle("CD4+\nT cells") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0) + theme_bw() 
  
  p_cd8 <- ggplot(data=data, aes(x=cluster_id, y=q_cd8)) + xlab("") + ylab("") + 
    ggtitle("CD8+\nT cells") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0) + theme_bw()
  
  p_nk16 <- ggplot(data=data, aes(x=cluster_id, y=q_nk16)) + xlab("") + ylab("") + 
    ggtitle("CD16hi\nNK cells") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0) + theme_bw()
  
  p_nk56 <- ggplot(data=data, aes(x=cluster_id, y=q_nk56)) + xlab("") + ylab("") + 
    ggtitle("CD56hi\nNK cells") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0) + theme_bw()
  
  p_nkt <- ggplot(data=data, aes(x=cluster_id, y=q_nkt)) + xlab("") + ylab("") + 
    ggtitle("NKT\ncells") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0) + theme_bw()
  
  p_neut <- ggplot(data=data, aes(x=cluster_id, y=q_neut)) + xlab("") + ylab("") + 
    ggtitle("Neutrophils") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0) + theme_bw() 
  
  p_eosin <- ggplot(data=data, aes(x=cluster_id, y=q_eosin)) + xlab("") + ylab("") + 
    ggtitle("Eosinophils") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0) + theme_bw()
  
  p_platelet <- ggplot(data=data, aes(x=cluster_id, y=q_platelet)) + xlab("") + ylab("") + 
    ggtitle("Platelets") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + expand_limits(y=0) + theme_bw()
  
  p_baso <- ggplot(data=data, aes(x=cluster_id, y=q_baso)) + xlab("") + ylab("") + 
    ggtitle("Basophils") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + expand_limits(y=0) + theme_bw()
  
  p_rbc <- ggplot(data=data, aes(x=cluster_id, y=q_rbc)) + xlab("") + ylab("") + 
    ggtitle("RBCs") + geom_violin(adjust=2)+ geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + expand_limits(y=0) + theme_bw()
  
  p_crp <- ggplot(data=data, aes(x=cluster_id, y=crp_macs)) + xlab("") + ylab("mg/L") + 
    ggtitle("CRP") + geom_violin(adjust=2)  + scale_y_log10() + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + theme_bw()
  
  p_il6 <- ggplot(data=data, aes(x=cluster_id, y=p_il_6_mean)) + xlab("") + ylab("pg/ml") + 
    ggtitle("Plasma IL-6") + geom_violin(adjust=2)  + scale_y_log10(labels = function(x) format(x, scientific = FALSE)) + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + 
    theme_bw()
  
  p_triglycerides <- ggplot(data=data, aes(x=cluster_id, y=triglycerides)) + xlab("") + ylab("mmol/L") + 
    ggtitle("Triglycerides") + geom_violin(adjust=2)  + scale_y_log10(labels = function(x) format(x, scientific = FALSE)) + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + 
    theme_bw()
  
  p_hamd17 <- ggplot(data=data, aes(x=cluster_id, y=ham_17_score)) + xlab("") + ylab("Total score") + 
    ggtitle("HAM-D\ntotal score") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + 
    expand_limits(y=c(0)) + theme_bw()
  
  p_chald <- ggplot(data=data, aes(x=cluster_id, y=chalder_fatigue_score)) + xlab("") + ylab("") + 
    ggtitle("Chalder fatigue\nrating") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + 
    expand_limits(y=c(0)) + theme_bw() 
  
  p_ctq <- ggplot(data=data, aes(x=cluster_id, y=ctq_total_score)) + xlab("") + ylab("Total score") + 
    ggtitle("Childhood\ntrauma score") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id), outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + expand_limits(y=c(0)) + 
    theme_bw() 
  
  # Life events questionnaire z-score  
  p_leq_z <- ggplot(data=data, aes(x=cluster_id, y=leq_z)) + xlab("") + ylab("Z-score") + 
    ggtitle("Recent\nstressors") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + theme_bw()
  
  p_bmi <- ggplot(data=data, aes(x=cluster_id, y=bmi_calculated)) + xlab("") + ylab("") + 
    ggtitle("BMI\n") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + 
    expand_limits(y=c(0)) + theme_bw() 
  
  p_age <- ggplot(data=data, aes(x=cluster_id, y=age_at_bloods)) + xlab("") + ylab("Years") + 
    ggtitle("Age\n") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE) + 
    expand_limits(y=c(0)) + theme_bw() 
  
  p_centre <- ggplot(data=data, aes(macs_flow_centre, fill = cluster_id, label=..count..)) + geom_bar(position = "fill") + ggtitle("Clinical\ncentre") + scale_fill_manual(values=clustercols) + 
    theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + xlab("") + ylab("Participants") +
    ggtitle("Study centre\n") + theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5)) + expand_limits(y=0) + 
    geom_text(stat="count",position="fill",vjust=1)
  
  p_bdi <- ggplot(data=data, aes(x=cluster_id, y=bdi_total_score)) + xlab("") + ylab("") + 
    ggtitle("BDI\nscore") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + 
    expand_limits(y=c(0)) + 
    theme_bw() 
  
  p_shaps <- ggplot(data=data, aes(x=cluster_id, y=shaps_score_calc)) + xlab("") + ylab("") + 
    ggtitle("SHAPS score\n(anhedonia)") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + 
    expand_limits(y=c(0)) + 
    theme_bw()
  
  p_stai_t <- ggplot(data=data, aes(x=cluster_id, y=t_anxiety_score)) + xlab("") + ylab("") + 
    ggtitle("STAI\n(trait anxiety)") + geom_violin(adjust=2)  + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + 
    expand_limits(y=c(0)) + 
    theme_bw()
  
  p_stai_s <- ggplot(data=data, aes(x=cluster_id, y=s_anxiety_score)) + xlab("") + ylab("") + 
    ggtitle("STAI\n(state anxiety)") + geom_violin(adjust=2) + geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + scale_fill_manual(values=clustercols) + guides(fill=FALSE)  + 
    expand_limits(y=c(0)) + 
    theme_bw()
  
  # Failed treatments
  p_failed75 <- ggplot(data=data, aes(x=cluster_id, y=antidepress_failed_75_less)) + xlab("") + ylab("# of ineffective treatments") + 
    ggtitle("Antidepressant\nnon-response") + geom_violin(adjust=2) + 
    geom_boxplot(aes(fill=cluster_id),outlier.shape=NA, notch=FALSE, width=0.2) + 
    scale_fill_manual(values=clustercols) + guides(fill=FALSE) + 
    expand_limits(y=c(0)) + guides(fill=FALSE) + scale_y_continuous(breaks=c(0,2,4,6,8)) +
    theme_bw()
  
  plist_numeric <- list(p_b=p_b,p_mc=p_mc,p_mi=p_mi,p_mnc=p_mnc,p_nk16=p_nk16,p_nk56=p_nk56,p_nkt=p_nkt,p_cd4=p_cd4,p_cd8=p_cd8,p_neut=p_neut,p_eosin=p_eosin,p_baso=p_baso,p_rbc=p_rbc,p_platelet=p_platelet,p_bmi=p_bmi,p_age=p_age,p_crp=p_crp,p_il6=p_il6,p_hamd17=p_hamd17,p_ctq=p_ctq,p_chald=p_chald,p_bdi=p_bdi,p_shaps=p_shaps,p_stai_t=p_stai_t,p_stai_s=p_stai_s,p_failed75=p_failed75, p_leq_z=p_leq_z, p_triglycerides=p_triglycerides)
  plist_categ <- list(p_centre,p_sex,p_infec,p_inflamm,p_tobacco)
  
  # Add statistics to plots  
  plist_numeric <- lapply(plist_numeric, function(x) x + 
                            geom_signif(comparisons = list(c(levels[1],levels[2])), margin_top = 0.15, test="wilcox.test", map_signif_level = c("***"=0.001, "**"=0.01, "*"=0.05, "ns"=1), textsize=5,vjust=1.2))
  
  # Make blank plot
  df <- data.frame()
  blank <- ggplot(df) + geom_blank(inherit.aes=FALSE) + theme_bw() + theme(panel.border = element_blank())
  
  
  graphs_input <- list(blank, plist_numeric$p_b,
                       plist_numeric$p_mc,
                       plist_numeric$p_mi,
                       plist_numeric$p_mnc,
                       plist_numeric$p_nk16,
                       plist_numeric$p_nk56,
                       plist_numeric$p_nkt,
                       plist_numeric$p_cd4,
                       plist_numeric$p_cd8,
                       plist_numeric$p_neut,
                       plist_numeric$p_eosin,
                       plist_numeric$p_baso,
                       plist_numeric$p_rbc,
                       plist_numeric$p_platelet)
  
  graphs_immuno <- list(plist_numeric$p_crp, plist_numeric$p_il6)
  
  graphs_interest <- list(blank, 
                          plist_numeric$p_hamd17,
                          plist_numeric$p_bdi,
                          plist_numeric$p_chald,
                          plist_numeric$p_shaps,
                          plist_numeric$p_stai_t,
                          plist_numeric$p_stai_s,
                          plist_numeric$p_failed75,
                          plist_numeric$p_ctq,
                          plist_numeric$p_leq_z) 
  
  graphs_figure <- list(plist_numeric$p_hamd17,
                        plist_numeric$p_bdi,
                        plist_numeric$p_chald,
                        plist_numeric$p_shaps,
                        plist_numeric$p_stai_t,
                        plist_numeric$p_stai_s,
                        plist_numeric$p_ctq,
                        plist_numeric$p_leq_z,
                        plist_numeric$p_failed75,
                        plist_numeric$p_bmi,
                        plist_numeric$p_age,
                        plist_numeric$p_crp,
                        plist_numeric$p_il6,
                        plist_numeric$p_triglycerides, blank, blank,  blank)
  
  graphs_confound <- list(plist_numeric$p_bmi,
                          plist_numeric$p_age,
                          p_centre + guides(fill=FALSE),
                          p_sex,p_infec,p_inflamm,p_tobacco, p_alcohol,p_cannabis,p_propAD)
  
  
  g <- cowplot::plot_grid(plotlist=graphs_input,nrow=3, align = "hv")   
  
  cowplot::save_plot(filename =  paste("apollo_clustering_force2_input_",reg,suffix,".pdf",sep=""), g, path=picpath, base_width=7, base_height=7)
  g <- cowplot::plot_grid(plotlist=graphs_interest,nrow=2, align = "h")   
  cowplot::save_plot(filename =  paste("apollo_clustering_force2_interest_",reg,suffix,".pdf",sep=""), g, path=picpath, base_width=7.5, base_height=5.3)
  g <- cowplot::plot_grid(plotlist=graphs_immuno,nrow=1, align = "h")   
  cowplot::save_plot(filename =  paste("apollo_clustering_force2_immuno_",reg,suffix,".pdf",sep=""), g, path=picpath, base_width=3.5, base_height=3)
  g <- cowplot::plot_grid(plotlist=graphs_confound,nrow=2, align = "h")   
  cowplot::save_plot(filename =  paste("apollo_clustering_force2_counfound_",reg,suffix,".pdf",sep=""), g, path=picpath, base_width=8, base_height=6)
  g <- cowplot::plot_grid(plotlist=graphs_figure,ncol=6, align = "hv")   
  cowplot::save_plot(filename =  paste("apollo_clustering_force2_numeric_",reg,suffix,".pdf",sep=""), g, path=picpath, base_width=9, base_height=7.5)
  
}

```

```{r}

mdd$cluster_id <- mdd$meta_two_way_resid
levels <- levels(mdd$cluster_id)

clustercols <- c("dark grey","red")

library(PMCMRplus) 
library(cowplot)

make_mini_cluster_plot_force2(data=mdd,basedir=here::here("res/"),picpath=here::here("pics/"), reg="reg",cluster_fx=choice_reg,choice_qset = choice_qset, cols=clustercols, levels=levels, suffix="_meta_two_way_resid")
```

Make the legend
```{r}

table(mdd$cluster_id) 

p <- ggplot(mdd,aes(x=cluster_id,y=ham_17_score, fill=cluster_id)) + geom_violin() + scale_fill_manual(values=clustercols, labels=c("Cluster 0, n=104","Cluster 1, n=102")) + theme(legend.title = element_blank())

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

p <- ggdraw() + draw_plot(get_legend(p)) 
ggsave(p, file=paste(here::here("pics/"),"apollo_cluster_legend_mdd_meta_two_way_resid.pdf",sep=""))

```

Now overlay clustering results on the PCA plot of the residual cell counts
```{r}

library(ggfortify)
library(ggrepel)
autoplot(pca.out, data=mdd, colour='meta_two_way_resid', label=FALSE, loadings=FALSE, loadings.label=FALSE, loadings.label.size=4, loadings.label.hjust=-0.1, loadings.label.repel=FALSE, frame = TRUE, frame.type="convex", frame.colour = 'meta_two_way_resid') + scale_color_manual(values = clustercols) + scale_fill_manual(values=clustercols)+ theme_bw() + guides(fill=FALSE,color=FALSE)
ggsave(filename = paste(here::here("pics/"),"apollo_pca_with_meta_twoway_resid.pdf",sep=""), width=4,height=4)

```

```{r}

sessionInfo()

```
